### 2021-11-04 paper/main.tex
add paper to repo

\documentclass{article}
\usepackage{caption}
\usepackage{wrapfig}
\usepackage[utf8]{inputenc} % allow utf-8 input
\usepackage{comment}
\usepackage[T1]{fontenc}    % use 8-bit T1 fonts
\usepackage{hyperref}       % hyperlinks
\usepackage{xurl}            % simple URL typesetting
\usepackage{booktabs}       % professional-quality tables
\usepackage{amsfonts}       % blackboard math symbols
\usepackage{nicefrac}       % compact symbols for 1/2, etc.
%

\usepackage{microtype}      % microtypography
\usepackage{todonotes}
\usepackage{longtable}
\usepackage[ruled,vlined]{algorithm2e}
\usepackage{amsmath}
\usepackage{utils/kky}
\usepackage{subcaption}
\usepackage{afterpage}
\usepackage{float}
\usepackage{ bbold }
\title{FastSRM: A fast, memory efficient and identifiable implementation of the
  shared response model}
\date{\today} % clear date

\begin{document}
\maketitle

\begin{abstract}
  The shared response model (SRM) provides a simple but effective framework to analyse
  fMRI data of subjects exposed to naturalistic stimuli.
  %
  However when the number
of subjects or runs is large, fitting the model requires a large amount of memory
and computational power, which limits its use in practice.
%
Furthermore, SRM is
not identifiable which makes the shared response difficult to interpret.
%

In this work, we implement an identifiable version of SRM and show on real data
that it improves the stability of the recovered shared response.
%
We then introduce the FastSRM that relies on a dimension reduction step that we
prove to be optimal in the sense that working with reduced data does not induce any change in the
algorithm trajectory.
%
We show experimentally using synthetic and real fMRI data
that FastSRM is considerably faster and more memory efficient
than competitive implementations.
%

The experiments performed in the paper are \emph{fully} reproducible: our
code available at \url{https://github.com/hugorichard/FastSRM} downloads the
data, runs the experiments and plot the figures.
%

\end{abstract}

\section{Introduction}
What makes the analysis of functional magnetic resonance imaging (fMRI) data
particularly difficult is that the number of features (voxels) is usually much
larger than the number of available samples (timeframes).
%
In such cases, dimension reduction methods are precious tools that allow to
reduce the computational burden of subsequent data analysis.
%
The shared response
model~\cite{chen2015reduced} provides a principled way to perform dimension
reduction of brain imaging data from subjects performing the same task and is
therefore a natural basis for common source
identification~\cite{richard2021model}~\cite{richard2020modeling},
classification~\cite{turek2018capturing}~\cite{chen2017shared}\cite{zhang2016searchlight}
or transfer learning~\cite{zhang2018transfer}.
%

However SRM has initially been designed to work within regions of interest using
only few subjects.
%
When using full brain data, computational costs become
high.
%
In addition, memory requirements are difficult to meet since the full dataset
needs to hold in memory.
%
Another issue of SRM is its non-identifiability (see the proof
in~\cite{richard2020modeling} Appendix D) which makes it challenging to
interpret the extracted shared components as they are defined up to a rotation.
%

In this work, we introduce FastSRM, an identifiable model that uses an optimal
atlas based decomposition to speed up the computations with provably no loss of
performance.
%
FastSRM gains identifiability by imposing that the covariance of the shared
components is diagonal (see ~\cite{richard2020modeling} Appendix D for a proof).
%
The gains in speed and memory usage are based on the intuition that a compressed
representation of the input should be good enough to find a suitable estimate of
the shared components as they live in reduced space.
%
It turns out that there
exists an optimal atlas that reduces the data in a such a way that the obtained
shared components (and other parameters) are the same as if they were computed
on full data.
%
We show that FastSRM brings several practical benefits: on real data, it produces more stable
estimates of the shared components than its non-identifiable counter-part and 
on synthetic as well as on real data, it is much faster and more memory
efficient than other implementations that do not make use of the optimal atlas.
%

Our python implementation is available at
\url{https://github.com/hugorichard/FastSRM}.
%
All our experiments are fully reproducible.
%
Scripts to download the data, run the experiments and plot the figures are
included.
%
A continuous integration pipeline runs the tests
automatically when any change to the core algorithm is made.
%

\section{Background: the shared response model (SRM)}
\label{sec:srm:review}
The shared response model~\cite{chen2015reduced} is a multi-view latent factor
model.
%
The data $\xb_1 \dots \xb_m$ are modeled as random vectors following the model:
\begin{align}
 &\xb_i = A_i \sbb + \nb_i \\
  &A_i^{\top}A_i = I_p
  \label{eq:model:srm}
\end{align}
where $\xb_i \in \RR^v$ is the data of view $i$, $A_i \in \RR^{p \times v}$ is the
mixing matrix of view $i$, $\nb_i$ is the noise of view $i$ and $\sbb \in \RR^p$ are the
shared components referred to as the \emph{shared response} in fMRI
applications.
%

The mixing matrices
$A_i$ are assumed to be orthogonal so that $A_i^{\top}A_i = I_p$.
%
However, in
general the matrix $A_i A_i^{\top}$ is different from identity.
%
The noise
$\nb_i$ is assumed to be Gaussian with covariance $\Sigma_i$ and independent
across views.
%
We assume the number of features $v$ to be much larger than the
number of components $p$: $v \gg p$.
%

The conceptual figure~\ref{fig:srm:conceptual_figure} illustrates an 
application of the shared response model to fMRI data.
%
The mixing
matrices are spatial topographies specific to each subjects while the shared
components give the common timecourses.
%
In~\cite{chen2015reduced, anderson2016enabling}, two versions of the shared
response model are introduced which we now present.
%

\begin{figure}
  \centering
  \includegraphics[scale=0.3]{figures/srm/conceptual_figure31.png}
  \caption{\textbf{Shared response model}: The raw fMRI data are modeled as a
    weighted combination of subject-specific spatial components with additive
    noise.
    %
    The weights are shared between subjects and constitute the shared response
    to the stimuli.
    %
  }
  \label{fig:srm:conceptual_figure}
\end{figure}

\subsection{Deterministic shared response model}
\label{sec:deterministicsrm}
Let us consider $n$ observations of $\xb_i$ and $\sbb$ that we stack into
matrices $X_i \in \RR^{v, n}$ and $S \in \RR^{p, n}$.
%

The deterministic shared response model sees both the mixing matrices $A_i$ and
the $n$ observations of the shared response $S$ as parameters to be
estimated.
%
The noise variance is fixed to a multiple of identity: $\forall i,
\Sigma_i=\sigma^2 I_v$ where $\sigma$ is an hyper-parameter to choose.
%

The model is optimized by maximizing the log-likelihood.
%

The likelihood is given by: $p(\xb) = \prod_i \Ncal(\xb_i; A_i \sbb, \sigma^2 I)$ and
therefore the empirical expected negative log-likelihood is given up to a constant independent of
$A_i$ and $S$ by:
\begin{align}
  \loss = \frac1{n} \sum_i \|A_i S - X_i \|^2 = \frac1{n} \big(\| S \|^2 -2 \dotp{ A_i S}{ X_i } + \| X_i \|^2 \big)
  \label{eq:detsrmloss}
\end{align}
$\loss$ is optimized by performing alternate minimization on $(A_1 \dots A_m)$
and $S$.
%
Note that the hyper-parameter $\sigma$ does not have an influence on
the loss and can therefore be safely ignored.
%

The gradient with respect to $S$ is given by $\sum_i A_i^{\top}(A_i S -
X_i) = \sum_i (S -
A_i^{\top} X_i)$
yielding the closed form updates:
\begin{equation}
  S \leftarrow  \frac1m \sum_i (A_i^{\top} X_i)
  \label{eq:srm:supdate}
\end{equation}

From \eqref{eq:detsrmloss}, minimizing $\loss$ with respect to $A_i$ is
equivalent to maximizing $\dotp{ A_i}{ X_i S^{\top} }$ and therefore we
have:
\begin{equation}
  A_i \leftarrow  \Pcal(\frac1{n}X_i S^{\top})
  \label{eq:detsrm:Aiupdate}
\end{equation}
where $\Pcal$ is the projection on the Stiefel manifold: $\Pcal(M) = M
(M^{\top}M)^{-\frac12}$.
%

The complexity of Deterministic SRM is in $\bigO(\mathrm{n_{iter}} mpvn)$ where
$n$ is the number of samples and $\mathrm{n_{iter}}$ the number of iterations.
%
We monitor the convergence by looking at the $\ell_{\infty}$ norm of the
gradient.
%
Note that we can monitor the gradient without any increase in complexity.
%
Indeed, after the updates with respect to each mixing matrix have been
carried out, only the gradient with respect to $S$ remains: $\sum_i
(S - A_i^{\top}\xb_i)$.
%
The algorithm is stopped when the
gradient falls below a chosen tolerance.
%

\subsection{Probabilistic SRM}
\label{sec:probabilisticsrm}
In Probabilistic SRM , $\Sigma_i=\sigma_i^2 I_v$ and the shared
components are assumed to be Gaussian $\sbb \sim \Ncal(0, \Sigma_s)$.
%

In~\cite{chen2015reduced} and~\cite{anderson2016enabling}, $\Sigma_s$ is only assumed to be definite positive.
%
As already highlighted in introduction, this causes the model to be
unidentifiable (see~\cite{richard2020modeling} Appendix D for a proof). Enforcing a diagonal $\Sigma_s$ ensures identifiability (provided the diagonal values are different).
%
So we assume here that $\Sigma_s$ is diagonal (and refer the interested reader
to~\cite{chen2015reduced} and~\cite{anderson2016enabling} for the original formulation of Probabilistic SRM
without the diagonal constraint)
%

The model is optimized via the expectation maximization algorithm.
%
Denoting $\VV[\sbb | \xb] = (\sum_i \frac1{\sigma_i^2} I +
\Sigma_s^{-1})^{-1}$ and $\EE[\sbb | \xb] = \VV[\sbb | \xb] \sum_i \frac1{\sigma_i^2}
A_i^{\top}\xb_i$, we have
\begin{align}
  p(\xb, \sbb) &= \prod_i \frac{\exp(-\frac{\|\xb_i - A_i \sbb \|^2}{2 \sigma_i^2})}{(2 \pi \sigma_i^{2v})^{\frac12}} \frac{\exp(-\frac12 \dotp{ \sbb }{ \Sigma_s^{-1} \sbb } )}{(2 \pi | \Sigma_s|)^{\frac12}} \\
               &= c_1 \exp(-\frac12 \left( \sum_i \frac1{\sigma_i^2}\|\xb_i\|^2 - 2  \dotp{ \sum_i \frac1{\sigma_i^2} A_i^{\top}\xb_i}{ \sbb } \right. \\& \left.+ \sum_i \frac1{\sigma_i^2} \| \sbb \|^2 + \dotp{ \sbb}{ \Sigma_s^{-1} \sbb }  \right)) \\
               &= c_2(\xb) \exp(-\frac12 \left( \dotp{  \sbb - \EE[\sbb }{ \xb], \VV[\sbb | \xb]^{-1} ( \sbb - \EE[\sbb | \xb])  } \right)) \label{probsrmcompleted}
\end{align}
where
\begin{align}
  c_1 = \frac1{(2 \pi \sigma_i^{2v})^{\frac12}}\frac1{(2 \pi |
  \Sigma_s|)^{\frac12}}
\end{align}
and
\begin{align}
  c_2(\xb) = c_1 \exp(-\frac12( \sum_i
  \frac1{\sigma_i^2}\|\xb_i\|^2 - \dotp{  \EE[\sbb }{ \xb], \VV[\sbb | \xb]^{-1}
  \EE[\sbb | \xb] }))
\end{align}
are independent of $\sbb$.
%
Therefore,
\begin{align}
  \sbb| \xb \sim \Ncal(\EE[\sbb | \xb], \VV[\sbb, \xb])
\end{align}

The negative expected completed log-likelihood is given by
\begin{align}
	\loss = \sum_i \frac12 v \log(\sigma_i^2) + \frac1{2 \sigma_i^2} \EE[\| \xb_i - A_i \sbb \|^2]
\end{align}
updates are therefore given by:
\begin{align}
&\sigma_i^2 \leftarrow \frac1{v} (\EE[\| \xb_i - A_i \EE[\sbb|\xb]\|^2] + \| \diag(\VV[\sbb | \xb]) \|^2) \\
  &A_i \leftarrow \Pcal(\EE[\xb_i \EE[\sbb|\xb]^{\top}]) \label{eq:psrm:Aiupdate} \\
  & \Sigma_s \leftarrow \VV[\sbb | \xb] + \EE[\EE[\sbb | \xb] \EE[\sbb | \xb]^{\top}]
\end{align}

It is useful to access the log-likelihood to check the implementation of the
algorithm and monitor the convergence.
%
From equation~\eqref{probsrmcompleted},
the likelihood is given by:
\begin{align}
  p(\xb) &= c_2(\xb) \int_{\sbb} \exp(-\frac12 \left( \dotp{  \sbb - \EE[\sbb }{ \xb], \VV[\sbb | \xb]^{-1} ( \sbb - \EE[\sbb | \xb])  } \right)) d\sbb \\
         &= c_2(\xb) (2 \pi |\VV[\sbb | \xb]|)^{\frac12}
\end{align}
replacing $c_2(\xb)$ by its expression and taking the log, the expected negative
log-likelihood is (up to constants) given by:
\begin{align}
  \EE[-\log(&p(\xb))] = \sum_i \frac{v}{2} \log(\sigma_i^2) + \frac12 \log(|\Sigma_s|) - \frac12 \log(|\VV[\sbb | \xb]|) \nonumber \\ &+ \sum_i
  \frac12 \frac1{\sigma_i^2} \EE[\|\xb_i\|^2] - \frac12 \EE[\dotp{  \EE[\sbb }{ \xb], \VV[\sbb | \xb]^{-1} \EE[\sbb | \xb] }]
\end{align}

The complexity of Probabilistic SRM is $\bigO(\mathrm{n_{iter}} mpvn)$, the same as in
Deterministic SRM.
%
We can monitor the convergence by looking at the log-likelihood decrease at each iteration
and stop the algorithm when the magnitude of the decrease is below some
tolerance.
%
The storage requirements of Deterministic or Probabilistic SRM are in
$\bigO(mvn)$ which simply means that the dataset needs to hold in memory.
%

\section{The FastSRM algorithm}

\subsection{Reducing the computational burden by use of atlases}
SRM algorithms use different set of parameters $\theta$ to represent the data.
%
In deterministic SRM $\theta = (A_i)_{i=1}^m, S$ where $(A_i)_{i=1}^m$ are the
mixing matrices and $S$ are the $n$ observations of the shared response $S$ while in probabilistic SRM $\theta
= (A_i)_{i=1}^m, \Sigma_s, (\sigma_i)_{i=1}^m$ where $(A_i)_{i=1}^m$ are the
mixing matrices, $(\sigma_i)_{i=1}^m$ the noise levels and $\Sigma_s$ the
components variance.
%

In fMRI, the classical approach used to reduce the data is to apply an atlas.
%
A deterministic atlas such as~\cite{bellec2010multi} is a parcellation of the
brain into $r$ regions.
%
Reducing an image using a deterministic atlas corresponds to
averaging the signal within each region of the atlas.
%
A probabilistic atlases such
as~\cite{dadi_fine-grain_2020} describes each region as a set of weights across
the full brain.
%
Therefore, the image reduction can be done with a matrix product.
%

In FastSRM we consider a set of view specific atlases $U_i \in \RR^{v \times r}$ such that
$U_i^{\top}U_i = I_r$ where $r$ is the number of regions in the atlas.
%
Data are reduced using $\zb_i = U_i^{\top} \xb_i$ and an SRM algorithm is applied
on data $\zb_i$ yielding parameters $\theta'$.
%
The figure~\ref{fig:srm:conceptual} illustrates this process.
%

Note that the parameters obtained with FastSRM $\theta'$ are different
from the parameters obtained with the corresponding SRM algorithm $\theta$ (the
unmixing matrices in $\theta'$ do not even have the same shape as the unmixing
matrices in $\theta$).
%
However, as we will see in the next section, there exists atlases such as the
correspondence between $\theta$ and
$\theta'$ can be made explicit.
%

\begin{figure}
  \centering
  \includegraphics[scale=0.34]{figures/srm/conceptual_figure2.pdf}
  \caption{\textbf{FastSRM algorithm} In step 1, data $\xb_i$ are projected onto an
    atlas $U_i$ that may depend on the subject $i$ (top).
    %
    In step 2 a SRM algorithm is applied on reduced data to compute the shared
    response.
    %
  }
  \label{fig:srm:conceptual}
\end{figure}

From a computational stand point, the dimension reduction provides
a large reduction in memory usage.
%
Indeed as the original data are seen only
once, it is no longer necessary to keep the full dataset in memory (we can load
data $X_i$ one after the other and similarly for the atlases $U_i$).
%
Therefore
the memory consumption is only in $\bigO(vn)$ (where $v$ is the number of voxels
and $n$ is the number of samples) which is lower than SRM by a factor of $m$,
the number of subjects.
%
The number of subjects is typically between $10$ and
$1000$.
%
This yields a practical benefit: on fMRI datasets with many subjects, one no
longer needs a large cluster to run the shared response model but only a modern
laptop.
%
Additionally, low memory consumption reduces the
risk of thrashing~\cite{denning1968thrashing}, a phenomenon that causes large
increase in computation time when the memory used is close to the total available
memory in the hardware.
%

After preprocessing, the reduced representation $\zb_i$ is used instead of the
original data $\xb_i$ yielding a time complexity of $\bigO(
\mathrm{T_{preprocessing}} + \mathrm{n_{iter}} mpnr)$.
%
Let us highlight that an experiment is often run
multiple times such as when cross validated results are needed.
%
In these cases,
the pre-processing is performed only once and the apparent complexity becomes
$\bigO(\mathrm{n_{iter}} mpnr)$ which is faster than SRM by
a factor of $\frac{v}{r}$.
%
The number of regions in large atlases is about $r=1000$ and in full brain data,
the number of voxels is about $300~000$ so that $\frac{v}{r}$ is typically about
$1000$.
%
It remains to show how to draw a correspondence between FastSRM and SRM which is
addressed in the following section.
%

\subsection{An optimal atlas}
\label{sec:optimal_atlas}
In principle, FastSRM can be used with any atlas.
%
However, in general, working with reduced data induces a loss of information
that can be minimized if the atlas is carefully chosen.
%
In any case, there is little hope to recover the parameters that would have been
obtained from SRM from the parameters of FastSRM.
%
Yet, we show that there exists an optimal atlas in the
sense that SRM and FastSRM yield the same results.
%

Let us consider $\xb_i = U_{\xb_i} \zb_i$ a PCA of $\xb_i$ with the maximum
number of components.
%
As the number of
samples $n$ is lower than the number of features, $U_{\xb_i} \in \RR^{v \times
  n}$ and $\zb_i \in \RR^{n}$.
%
We also have $U_{\xb_i}^{\top}U_{\xb_i} = I$.
%
Therefore $U_{\xb_i}$ constitutes a possible choice of subject specific atlas.
%
As the next property shows, $U_{\xb_i}$ is an optimal atlas for deterministic
FastSRM.
%

\begin{proposition}[Optimal atlas for deterministic FastSRM]
  Let $(A_i)_i, S$ be the solution obtained by deterministic SRM on data
  $(X_i)_i$ and $(A_i')_i, S'$ the solution obtained by deterministic FastSRM on
  data $(X_i)_i$ using atlases $(U_{X_i})_i$ where $X_i = U_{\xb_i} Z_i$ is a PCA
  of $X_i$.
  %
  Then $A_i = U_{X_i}A_i'$ and $S = S'$.
  %

\label{prop:optimaldetsrm}
\end{proposition}
\begin{proof}
Updates of the mixing matrices $A_i$ in deterministic SRM
equation~\eqref{eq:detsrm:Aiupdate} can be written:
\begin{align}
  A_i &\leftarrow \Pcal(\frac1{n}X_i S^{\top})= U_{X_i}\Pcal(\frac1{n}Z_i S^{\top})
  \label{eq:fastdetsrm:Aiupdate}
\end{align}
where $\Pcal$ is the projection on the Stiefel manifold: $\Pcal(M) = M
(M^{\top}M)^{-\frac12}$.
%

Therefore we can look for $A_i$ as $A_i = U_{X_i} \tilde{A}_i$.
%
We can show that $\tilde{A}_i$ is orthogonal.
%
Indeed,
\begin{align}
  &A_i^{\top} A_i = I_p \\
  & \implies \tilde{A}_i^{\top}U_{X_i}^{\top} U_{X_i} \tilde{A}_i = I_p \\
  & \implies \tilde{A}_i^{\top} \tilde{A}_i = I_p
\end{align}

Then, we use the fact that
\begin{align}
  \|X_i - A_i S \|^2 = \| U_{X_i}Z_i - U_{X_i}\tilde{A}_i S\|^2 = \| Z_i - \tilde{A}_i S \|^2
  \label{eq:equality:xy}
\end{align}
so that $A_i' = \tilde{A}_i$.
%

Therefore, the solution of deterministic SRM on data $(\zb_i)_{i=1}^m$ and
$(\xb_i)_{i=1}^m$ are linked by the change of parameters $A_i = U_{\xb_i}A_i'$ and
$S = S'$.
%
This concludes the proof.
%

\end{proof}

In the case of probabilistic SRM we can obtain very similar results.
%
However the
algorithm applied on reduced data need to be slightly modified.
%

We call probSRM($\psi$) the probabilistic SRM algorithm modified such that
updates
\begin{align}
\sigma_i^2 \leftarrow \frac1{v} (\EE[\| \xb_i - A_i \EE[\sbb|\xb]\|^2] + \| \diag(\VV[\sbb | \xb]) \|^2)
\end{align}
are replaced by updates
\begin{align}
  \sigma_i^2 \leftarrow \frac1{\psi} (\EE[\| \xb_i - A_i \EE[\sbb|\xb]\|^2] + \| \diag(\VV[\sbb | \xb]) \|^2)
\end{align}

We have the following result:
\begin{proposition}[Optimal atlas for probabilistic FastSRM]
  Let $(A_i)_i, \sigma_i, \Sigma_s$ be the solution obtained by probabilistic SRM on data
  $\xb_i$ and $(A_i')_i, \sigma_i', \Sigma_s'$ the solution obtained by ProbSRM($v$) on
  data $\zb_i = U_{\xb_i}^{\top} \xb_i$.
  %
  Then $A_i = U_{\xb_i}A_i'$, $\sigma_i =
  \sigma_i'$ and $\Sigma_s = \Sigma_s'$.
  %

  \label{prop:optimalprobsrm}
\end{proposition}
\begin{proof}
  Updates of the mixing matrices $A_i$ in probabilistic SRM
  equation~\eqref{eq:psrm:Aiupdate} can be written:
  \begin{align}
    A_i &\leftarrow U_{\xb_i}\Pcal(\EE[\zb_i \EE[\sbb| \xb_i]^T])
    \label{eq:fastprobsrm:Aiupdate}
  \end{align}
  so we can look for $A_i$ as $A_i = U_{\xb_i} \tilde{A}_i$ and, as in the
  deterministic case, $\tilde{A}_i$ is orthogonal.
  %
  Therefore equality~\eqref{eq:equality:xy} holds.
  %

  Then we consider the expected negative log-likelihood of probabilistic srm:
  \begin{align}
    \loss &= \sum_i \frac12 v\log(\sigma_i^2) + \frac12 \log(|\Sigma_s|) + \EE[\int_{\sbb} \sum_i \frac1{2 \sigma_i^{2}} \|\xb_i - A_i \sbb \|^2 \nonumber \\& \enspace \enspace \enspace \enspace+ \frac12 \dotp{ \sbb }{ \Sigma_s^{-1} \sbb }  d\sbb] \\
          &= \sum_i \frac12 v \log(\sigma_i^2) + \frac12 \log(|\Sigma_s|) + \EE[\int_{\sbb} \sum_i \frac1{2 \sigma_i^{2}}\|\zb_i - \tilde{A}_i \sbb \|^2 \nonumber \\& \enspace \enspace \enspace \enspace+ \frac12 \dotp{ \sbb }{ \Sigma_s^{-1} \sbb }  d\sbb]
  \end{align}
  where we use equality~\eqref{eq:equality:xy}.
  %

  Optimizing the log-likelihood via expectation maximization yields the exact
  same updates as probabilistic SRM on data $\zb_i$
  except that updates
  \begin{align}
    \sigma_i^2 \leftarrow \frac1{t} (\EE[\| \zb_i - \tilde{A}_i \EE[\sbb|\zb]\|^2] + \| \diag(\VV[\sbb | \zb]) \|^2)
  \end{align}
  are replaced by updates
  \begin{align}
    \sigma_i^2 \leftarrow \frac1{v} (\EE[\| \zb_i - \tilde{A}_i \EE[\sbb|\zb]\|^2] + \| \diag(\VV[\sbb | \zb]) \|^2)
  \end{align}
  so that $\tilde{A}_i = A'_i$.
  %

  Therefore, the updates in both algorithms are linked  by $A_i = U_{\xb_i}A_i'$,
  $\sigma_i' =
  \sigma_i$ and $\Sigma_s'  = \Sigma_s$.
  %

  This concludes the proof.
  %

\end{proof}

Proposition~\ref{prop:optimaldetsrm} and Proposition~\ref{prop:optimalprobsrm} show that
no information is lost by replacing $\xb_i \in \RR^v$ by its reduced
representation $\zb_i \in \RR^n$.
%
A key observation is that Proposition~\ref{prop:optimaldetsrm} and
Proposition~\ref{prop:optimalprobsrm} hold whether or
not the model for deterministic (respectively probabilistic) SRM is indeed the
generative model of the data.
%

A complexity analysis shows that finding the optimal atlas becomes the limiting
step of the pipeline.
%
Even with fast implementations, the subject specific PCA
is costly.
%
However FastSRM only works on $\zb_i$ so we do not need to know the value of
$U_{\xb_i}$.
%
In practice, we observe data $X_i \in \RR^{v \times n}$ and we want to get $Z_i
\in \RR^{n \times n}$ such that $X_i = U_{\xb_i} Z_i$.
%
This can be done by performing an
SVD of $X_i^{\top} X_i$ yielding $X_i^{\top}X_i= V_i D_i V_i^{\top}$ and setting
$Z_i = D_i^{\frac12} V_i^{\top}$.
%
Although computing the product $X_i^{\top} X_i$ has time complexity
$\bigO(vt^2)$ (the same as a PCA) the constant in the $\bigO$ is exactly one so
this operation costs a lot less than the PCA
on full data.
%
When estimates of the mixing matrices are needed, they can be obtained by
applying equation~\eqref{eq:fastdetsrm:Aiupdate} in the deterministic SRM case and
equation~\eqref{eq:fastprobsrm:Aiupdate} in the probabilistic SRM case which only costs
$\bigO(mvp^2)$.
%
In practice the cost of the matrix products $X_i^{\top} X_i$ is often still the
limiting step of the pipeline (this depends on the number of iterations) but as
we show in the next chapter, it is much more efficient than performing SRM on
the full data.
%
Note than if memory allows it, these matrix products can be
computed in parallel.
%

\section{Related work}
The implementation of SRM proposed in~\cite{chen2015reduced} is quadratic in the
number of features which prevent its application to full brain data.
In~\cite{anderson2016enabling}, the authors address this issue by using the
inversion lemma to remove the quadratic dependence. Their implementation is now the most widely used one.
In our work, we propose to reduce further the computation time and memory usage
compared to the implementation in~\cite{anderson2016enabling} by the use of
optimal atlases.

Other popular methods used to speed up SRM are searchlight~\cite{zhang2016searchlight} or piecewise~\cite{bazeille2021empirical}
approaches. While these methods are efficient, they do not optimize the same
objective as the original SRM algorithm and are arguably less principled since the
searchlights or pieces are chosen a priori.

We see SRM as a well principled formulation of the hyperalignment
method~\cite{haxby2011common}.
%
Many methods exist to perform hyperalignment: deep
hyperalignment~\cite{yousefnezhad2017deep} , robust
SRM~\cite{turek2018capturing} , multi-view canonical correlation
analysis~\cite{li2009joint}, ShICA~\cite{richard2021model},
MultiViewICA~\cite{richard2020modeling} optimal
transport~\cite{bazeille2019local} and many more.
%
In this work, we do not claim that FastSRM is the best method. We only see that
SRM is widely used and aim to provide a faster, more memory efficient and
identifiable implementation.  In our view FastSRM should be seen as a plugin
replacement for SRM.

\section{Experiments}
We make several experiments on both synthetic and real data.
%
We used nilearn~\cite{abraham2014machine} for fMRI data preprocessing,
Brainiak~\cite{kumar2020brainiak2} for the non-identifiable version of SRM that
implements the work of~\cite{chen2015reduced} and~\cite{anderson2016enabling}, Numpy~\cite{harris2020array} for
array processing, Scipy~\cite{2020SciPy-NMeth} its implementation of the
Hungarian algorithm, Matplotlib~\cite{hunter2007matplotlib} for plotting and
Sklearn~\cite{pedregosa2011scikit} for machine learning pipelines.

\subsection{Comparing Fitting time and performance of FastSRM and
  SRM on synthetic data}
We generate synthetic data $\xb_i$ according to the model of probabilistic SRM.
%
The parameters $\sigma_i$, $A_i$ and $\Sigma_s$ are generated randomly.
%
We sample the value of the subject specific noise level from a normal
distribution: $\sigma_i \sim \Ncal(0, 0.1)$.
%
The mixing matrices $A_i$
are obtained by sampling their coefficient from a standardized normal
distribution.
%
Lastly, the covariance of the shared response $\Sigma_s$ is diagonal and the
diagonal values are obtained by sampling from a Dirichlet distribution with
parameter $(1 \dots 1)$.
%
We set the number of voxels to $v=125~000$, the number of subjects to $m=10$ and
the number of components to $p=50$.
%
We generate $n=1000$ samples.
%

We benchmark deterministic SRM, probabilistic SRM and their FastSRM counterparts in terms of fitting time and performance.
%
Note that in this section, the identifiable implementation of deterministic SRM
and probabilistic SRM described in section~\ref{sec:probabilisticsrm}
and~\ref{sec:deterministicsrm} is used. 
%
Algorithms are designated by the atlas they use and therefore SRM algorithms are refered to
as \emph{None} because no atlas is used and FastSRM algorithms will have the
label \emph{Optimal}.
%
Note that it would be possible to use FastSRM with sub-optimal
atlases (there exists a wide variety of atlases
available~\cite{schaefer2017local, bellec2010multi, mensch2018extracting}) but
without any guarantees that the performance are the same as SRM.
%

We use a number of iterations between 1 and 100 and report the performance,
fitting time and a measure of convergence.
%
In FastSRM, we do not compute the
unmixing matrices but only the shared response.
%
We measure the performance of an algorithm by computing the error between the true component $S \in \RR^{p \times n}$ and
the predicted component $\hat{S} \in \RR^{p \times n}$ using the quantity:
\begin{align}
\mathrm{mse}(\hat{S}, S) = min_{A \in \RR^{p \times p}} \|A \hat{S} - S \|^2_F =  \|S
\hat{S}^{\dagger}\hat{S} - S \|^2_F
\end{align}
%
This way of measuring errors is
insensitive to the indeterminacies in DetSRM.
%
We measure the fitting time in seconds.
%
Lastly, we measure convergence by computing the gradient $\ell_{\infty}$ norm in
case of DetSRM given by $\max(\mathrm{abs}(G))$ where $G$ is the gradient and
use the distance between consecutive values of the loss for ProbSRM.
%
Results are plotted in Figure~\ref{fig:srm:synthetic_gradient}.
%

We empirically see that the optimal approach is equivalent to using no atlas in
terms of performance.
%
This is predicted by the theory in section~\ref{sec:optimal_atlas} where we demonstrate that these two algorithms yield exactly the same output
from the same input.
%
% %
% Among approximate methods, the projection
% method yields the best results but are quite far away from optimal results.
%
% %
In general probabilistic methods give much better results than their deterministic
counterpart.
%
This shows the superiority of likelihood based methods.
%
In terms of fitting time, FastSRM is about a thousand time faster than SRM after
100 iterations.
%
When no atlas is used, the number
of iterations has a very strong impact on performance while it has a small impact
when the optimal atlas is used.
%
Lastly, looking at the convergence curves, we see that even after $100$ iterations, algorithms did
not fully converge.
%
This means that in practice a much larger number of
iterations is needed which would yield an even higher difference in fitting time
between methods using no atlas and methods using the optimal atlas.
%

\begin{figure}
  \centering
  \includegraphics[width=\textwidth]{figures/srm/synthetic_gradient.pdf}
  \caption{\textbf{Benchmark of SRM algorithms on synthetic data: } Performance,
    fitting time and convergence of SRM algorithms in the deterministic (left)
    or probabilistic (right) case.
    %
    As expected, when optimal atlases are used,
    the performance is the same as if no atlas is used but the fitting time is
    much lower.
    %
    This is even more pronounced when the number of iterations is
    high (and looking at convergence curves, we see that more iterations could
    be performed to be closer to a stationary point).
    %
  }
  \label{fig:srm:synthetic_gradient}
\end{figure}

\subsection{Experiment on fMRI data: identifiability increases stability}
We evaluate the performance of the different SRM implementations on the
\emph{Sherlock} datasets where fMRI of 17 participants watching "Sherlock" BBC
TV show (episode 1) is performed.
%
% 
These data are downloaded from
\url{http://arks.princeton.edu/ark:/88435/dsp01nz8062179}.
%
% 
Data were acquired using a 3T scanner with an isotropic spatial resolution of
3~mm.
%
% 
More information including the preprocessing pipeline is available in
\cite{sherlock}.
%
% 
Subject 5 is removed because of missing data, leaving us with 16 participants.
%
% 
Although SHERLOCK data contains originally only 1 run, we split it into 4 blocks
of 395 timeframes and one block of 396 timeframes for the needs of our
experiments.
%

We first show that identifiability is a desirable property as it increases
stability.
%
Then we show that our FastSRM algorithm (that works with optimal atlases)
matches the performance of SRM (that works on full data) but use much less
computational resources.
%

\subsection{Identifiability increases stability}
Assuming that the data follow the model, identifiability ensures that the
correct parameters can be recovered without ambiguity.
%
In real fMRI data, the model cannot be expected to hold perfectly, but we can
hope for greater stability in the parameters recovered than if an unidentifiable
model is used.
%

To measure the stability of the common components obtained from the Sherlock
dataset, we divide the subjects into two roughly equal groups and extract the
common components in each group.
%
The components are then matched using the Hungarian algorithm and the stability
measure is obtained from the average correlation of the matched components.
%
The procedure is repeated 9 times using the Brainiak implementation (that is not
identifiable) and our
FastSRM implementation.
%
We plot the results as a scatter plot in Figure~\ref{exp:identifiability}, where, for each repetition , the x-axis indicates the stability measure obtained with Brainiak's implementation
(not identifiable since the shared components covariance is unconstrained) and the y-axis the stability measure
obtained with FastSRM (identifiable since the shared components covariance is diagonal).
%
We see that introducing a diagonal source covariance improves stability.
%

\begin{figure}
\begin{minipage}{0.38\linewidth}
      \includegraphics[width=\textwidth]{figures/srm/identifiability.pdf}
    \end{minipage}
    \hfill
    \begin{minipage}{0.6\linewidth}
        \caption{\label{exp:identifiability}\textbf{Identifiability increases
            stability: } We first divide the subjects
          into two groups and extract the common components in each group.
          %
          The components of the two groups are then matched using the Hungarian
          algorithm and the stability index is determined by the average
          correlation of the matched components.
          %
          The procedure is repeated 9 times with the Brainiak implementation
          (not identifiable since the shared components covariance is unconstrained) and our FastSRM implementation
          (identifiable since the shared components covariance is diagonal).
          %
        }
    \end{minipage}
\end{figure}

\subsection{Comparing fitting time, memory usage and performance on a
  timesegment matching experiment}
\label{sec:timesegment_expe}
\label{timesegment_expe}
The timesegment matching experiment is first introduced
in~\cite{chen2015reduced}.
%
In a nutshell, the time-segment matching accuracy measures the similarity
between two multivariate time-series by trying to localize a time-segment in one
time-series by correlation with the other.
%
In the context of movie watching, this measure has a lot of sense: if we split
the movies in scenes and compute a representation per scene and per subject, it makes
sense to assume that different subjects watching the movie would still have
closer representation of the same scenes than of different scenes.
%
This explains why timesegment matching is a standard evaluation of SRM-like methods also used in  \cite{guntupalli2018computational}, \cite{Nastase741975} or
\cite{zhang2016searchlight}.
%

We now describe more precisely the experimental design.
%
We split the runs into a train and test set.
%
After fitting the model on the
training set, we apply the unmixing matrices $W_i=A_i^{-1}$ of each subject on
the test set yielding individual components matrices.
%
We estimate the shared responses by averaging the individual components of each
subjects but one.
%
We select a target time-segment (9 consecutive timeframes) in the shared
responses and try to localize the corresponding time segment in the components
of the left-out subject using a maximum-correlation classifier.
%

The time-segment is said to be
correctly classified if the correlation between the sample and target
time-segment is higher than with any other time-segment (partially overlapping
time windows are excluded).
%

% 
We use 5-Fold cross-validation across runs: the training set contains 80\% of
the runs and the test set 20\%, and repeat the experiment using all possible
choices for left-out subjects.
%
% 
The mean accuracy is reported in Figure~\ref{fig:srm:timesegment} (bottom).
%
When the
optimal atlas is used, the accuracy is the same as when no atlas is used but
the fitting time is reduced by a factor 10 to 100 and so is the memory usage.
%

%

\begin{figure}
  \centering
  \includegraphics[width=0.8\textwidth]{figures/srm/timesegment_matching_sherlock.pdf}
  \caption{\textbf{Benchmark of SRM algorithms on fMRI data} (top) Timesegment
    matching accuracy (middle) Fitting time (bottom) Memory usage. When the
    optimal atlas is used, the accuracy is the same as when no atlas is used but
    the fitting time is reduced by a factor 10 to 100 and so is the memory usage.
    %
  }
  \label{fig:srm:timesegment}
\end{figure}

\section{Conclusion}
As studies using naturalistic stimuli will tend to become more common and large
within and across subjects, we need scalable models in terms of computation time
and memory usage.
%
This is what FastSRM provides.

%
FastSRM is an efficient implementation of SRM
that uses optimal atlases to speed up computations and reduce memory requirements
with provably no loss of performance.

We show on synthetic and real data that FastSRM is much faster and more memory
efficient that implementations not using the optimal atlas.
Furthermore, FastSRM is identifiable. On real data, we show that compared to
Brainiak's implementation (which is not identifiable), FastSRM provides more
stable estimates of the shared components.

%
FastSRM inherits from all the weaknesses of SRM including the fact that
mixing matrices are assumed to be orthogonal which is rather unrealistic. It
remains to be seen whether dimension reduction and learning of the shared
components could be done jointly and efficiently without assuming orthogonal
mixing matrices.
%
Our optimal atlas provide the intuition that SRM-like methods do not need the
full data to provide accurate estimates of shared components. We believe such
insights may guide the design of future methods.

\bibliographystyle{plain}
\bibliography{utils/biblio}
\end{document}